\singlespacing \vspace{-1.6cm} \LARGE Covariation among reproductive traits in flowering plants determine interactions with floral visitors

### 2022-06-09 Manuscript/Lanuza_et_al_2021.Rmd
edit

Flowering plants have an astonishing diversity of floral structures [@barrett2002; @schiestl2013] that shape plant-pollinator associations [@fenster2004; @dellinger2020]. However, not all reproductive trait combinations are possible due to evolutionary and ecological constraints [@stearns1989]. Despite the recent advances in the theoretical and empirical understanding of the macroecological correlations between plant reproductive traits [@salguero2016; @friedman2020; @roddy2021], their study is often limited to a handful number of reproductive traits that are rarely studied jointly. Thus, in order to progress towards a comprehensive understanding of the plant reproductive spectrum of trait variation as done recently for other vegetative and physiological plant traits [@wright2004; @chave2009; @diaz2016; @onoda2017; @laughlin2021], there is a need to acquire a multitrait perspective with broad geographical coverage. Importantly, the characterization of the reproductive trait covariation patterns can help to further understand the different plant ecological strategies [@agrawal2020] and improve our knowledge on plant-pollinator associations [@roddy2021].

@@

This study investigates the reproductive spectrum of flowering plants from plant-pollinator networks and shows that plants have clear positive and negative correlations between reproductive traits that determine interactions with floral visitors at a global scale. We found that there are three reproductive axes that capture over 70% of the reproductive trait variation: (i) the flower number - flower size axis, (ii) the pollinator dependence axis and (iii) the style length - pollinator dependence axis. These reproductive axes helped partly explain the presence of floral visitor interactions, but not their number of visits. However, floral visitor guilds formed distinct relationships with the main axes of trait variation. Finally, we found that the plant species network roles were best explained by plant size and floral reward related traits.

>>>>

Flowering plants have an astonishing diversity of floral structures [@barrett2002; @schiestl2013] that shape plant-pollinator associations [@fenster2004; @dellinger2020]. However, not all reproductive trait combinations are possible due to evolutionary and ecological constraints [@stearns1989]. Despite the recent advances in the theoretical and empirical understanding of the macroecological correlations between plant reproductive traits [@salguero2016; @friedman2020; @paterno2020; @roddy2021], their study is often limited to a handful number of reproductive traits that are rarely studied jointly. Thus, in order to progress towards a comprehensive understanding of the plant reproductive spectrum of trait variation as done recently for other vegetative and physiological plant traits [@wright2004; @chave2009; @diaz2016; @onoda2017; @laughlin2021], there is a need to acquire a multitrait perspective with broad geographical coverage. Importantly, the characterization of the reproductive trait covariation patterns can help to further understand the different plant ecological strategies [@agrawal2020] and improve our knowledge on plant-pollinator associations [@roddy2021].

@@

This study investigates the reproductive spectrum of flowering plants and shows that plants have clear positive and negative correlations between reproductive traits that determine interactions with floral visitors at a global scale. Despite the enormous variability of plant reproductive structures, three axes were able to capture over 70% of reproductive trait variation. The first axis was the flower number - flower size and described the negative association between flower number and flower size. Although this negative correlation has received substantial attention [worley2000; @sargent2007; @kettle2011], it has only been investigated across a few hundred of Angiosperm species and without a multitrait perspective. Here, we provide support to this evidence and show that also key reproductive traits to plant reproduction like ovule number and style length are positively correlated with flower size and negatively correlated with flower number. The second axis organized the species by the degree of pollinator dependence where species with higher pollinator dependence showed greater floral display.

As expected... quizas hilar aqui con un par de refs para que quede menos results

it has been investigated empirically only just with few hundreds of Angiosperm species and without a multitrait perspective. The second axis of reproductive trait variation was the pollinator dependence axis and showed the variation

main covariation pattern of the reproductive trait space was the negative correlation between flower number and flower size that is a widely described association across [@sargent2007; @kettle2011] and within [@delph2004] Angiosperms species. Our study explore this association beyond a two-trait correlation and shows that other fundamental traits to plant reproduction are also negatively correlated to flower number like ovule number and style length.

The second axis that explained most variation was the pollinator dependence one

We found that there are three reproductive axes that capture over 70% of the reproductive trait variation: (i) the flower number - flower size axis, (ii) the pollinator dependence axis and (iii) the style length - pollinator dependence axis. These reproductive axes helped partly explain the presence of floral visitor interactions, but not their number of visits. However, floral visitor guilds formed distinct relationships with the main axes of trait variation. Finally, we found that the plant species network roles were best explained by plant size and floral reward related traits.

### 2022-06-09 Manuscript/Lanuza_et_al_2021.tex
edit

Flowering plants have an astonishing diversity of floral structures (\protect\hyperlink{ref-barrett2002}{Barrett, 2002}; \protect\hyperlink{ref-schiestl2013}{Schiestl \& Johnson, 2013}) that shape plant-pollinator associations (\protect\hyperlink{ref-dellinger2020}{Dellinger, 2020}; \protect\hyperlink{ref-fenster2004}{Fenster et al., 2004}). However, not all reproductive trait combinations are possible due to evolutionary and ecological constraints (\protect\hyperlink{ref-stearns1989}{Stearns, 1989}). Despite the recent advances in the theoretical and empirical understanding of the macroecological correlations between plant reproductive traits (\protect\hyperlink{ref-friedman2020}{Friedman, 2020}; \protect\hyperlink{ref-roddy2021}{Roddy et al., 2021}; \protect\hyperlink{ref-salguero2016}{Salguero-Gómez et al., 2016}), their study is often limited to a handful number of reproductive traits that are rarely studied jointly. Thus, in order to progress towards a comprehensive understanding of the plant reproductive spectrum of trait variation as done recently for other vegetative and physiological plant traits (\protect\hyperlink{ref-chave2009}{Chave et al., 2009}; \protect\hyperlink{ref-diaz2016}{Díaz et al., 2016}; \protect\hyperlink{ref-laughlin2021}{Laughlin et al., 2021}; \protect\hyperlink{ref-onoda2017}{Onoda et al., 2017}; \protect\hyperlink{ref-wright2004}{Wright et al., 2004}), there is a need to acquire a multitrait perspective with broad geographical coverage. Importantly, the characterization of the reproductive trait covariation patterns can help to further understand the different plant ecological strategies (\protect\hyperlink{ref-agrawal2020}{Agrawal, 2020}) and improve our knowledge on plant-pollinator associations (\protect\hyperlink{ref-roddy2021}{Roddy et al., 2021}).

@@

This study investigates the reproductive spectrum of flowering plants from plant-pollinator networks and shows that plants have clear positive and negative correlations between reproductive traits that determine interactions with floral visitors at a global scale. We found that there are three reproductive axes that capture over 70\% of the reproductive trait variation: (i) the flower number - flower size axis, (ii) the pollinator dependence axis and (iii) the style length - pollinator dependence axis. These reproductive axes helped partly explain the presence of floral visitor interactions, but not their number of visits. However, floral visitor guilds formed distinct relationships with the main axes of trait variation. Finally, we found that the plant species network roles were best explained by plant size and floral reward related traits.

@@

@@

>>>>

Flowering plants have an astonishing diversity of floral structures (\protect\hyperlink{ref-barrett2002}{Barrett, 2002}; \protect\hyperlink{ref-schiestl2013}{Schiestl \& Johnson, 2013}) that shape plant-pollinator associations (\protect\hyperlink{ref-dellinger2020}{Dellinger, 2020}; \protect\hyperlink{ref-fenster2004}{Fenster et al., 2004}). However, not all reproductive trait combinations are possible due to evolutionary and ecological constraints (\protect\hyperlink{ref-stearns1989}{Stearns, 1989}). Despite the recent advances in the theoretical and empirical understanding of the macroecological correlations between plant reproductive traits (\protect\hyperlink{ref-friedman2020}{Friedman, 2020}; \protect\hyperlink{ref-paterno2020}{Paterno et al., 2020}; \protect\hyperlink{ref-roddy2021}{Roddy et al., 2021}; \protect\hyperlink{ref-salguero2016}{Salguero-Gómez et al., 2016}), their study is often limited to a handful number of reproductive traits that are rarely studied jointly. Thus, in order to progress towards a comprehensive understanding of the plant reproductive spectrum of trait variation as done recently for other vegetative and physiological plant traits (\protect\hyperlink{ref-chave2009}{Chave et al., 2009}; \protect\hyperlink{ref-diaz2016}{Díaz et al., 2016}; \protect\hyperlink{ref-laughlin2021}{Laughlin et al., 2021}; \protect\hyperlink{ref-onoda2017}{Onoda et al., 2017}; \protect\hyperlink{ref-wright2004}{Wright et al., 2004}), there is a need to acquire a multitrait perspective with broad geographical coverage. Importantly, the characterization of the reproductive trait covariation patterns can help to further understand the different plant ecological strategies (\protect\hyperlink{ref-agrawal2020}{Agrawal, 2020}) and improve our knowledge on plant-pollinator associations (\protect\hyperlink{ref-roddy2021}{Roddy et al., 2021}).

@@

This study investigates the reproductive spectrum of flowering plants and shows that plants have clear positive and negative correlations between reproductive traits that determine interactions with floral visitors at a global scale. Despite the enormous variability of plant reproductive structures, three axes were able to capture over 70\% of reproductive trait variation. One of the major covariation patterns of the reproductive trait space was the negative correlation between flower number and flower size that is a widely described association across (\protect\hyperlink{ref-kettle2011}{Kettle et al., 2011}; \protect\hyperlink{ref-sargent2007}{Sargent et al., 2007}) and within (\protect\hyperlink{ref-delph2004}{Delph et al., 2004}) Angiosperms species. Our study explore this association beyond a two-trait correlation and shows that other fundamental traits to plant reproduction are also negatively correlated to flower number like ovule number and style length.

The second axis that explained most variation was the pollinator dependence one

We found that there are three reproductive axes that capture over 70\% of the reproductive trait variation: (i) the flower number - flower size axis, (ii) the pollinator dependence axis and (iii) the style length - pollinator dependence axis. These reproductive axes helped partly explain the presence of floral visitor interactions, but not their number of visits. However, floral visitor guilds formed distinct relationships with the main axes of trait variation. Finally, we found that the plant species network roles were best explained by plant size and floral reward related traits.

@@

\leavevmode\vadjust pre{\hypertarget{ref-delph2004}{}}%
Delph, L. F., Gehring, J. L., Frey, F. M., Arntz, A. M., \& Levri, M. (2004). Genetic constraints on floral evolution in a sexually dimorphic plant revealed by artificial selection. \emph{Evolution}, \emph{58}(9), 1936--1946.

@@

\leavevmode\vadjust pre{\hypertarget{ref-paterno2020}{}}%
Paterno, G. B., Silveira, C. L., Kollmann, J., Westoby, M., \& Fonseca, C. R. (2020). The maleness of larger angiosperm flowers. \emph{Proceedings of the National Academy of Sciences}, \emph{117}(20), 10921--10926.
